\documentclass[11pt]{article}
    %	options include 12pt or 11pt or 10pt
    %	classes include article, report, book, letter, thesis

    \usepackage{amsmath}
    \usepackage{array}
    \setlength\extrarowheight{2pt}
    \usepackage{graphicx}
    \graphicspath{ {/home/shanedrafahl/coms331/hw0} }

    \title{HW1}
    \author{Shane Drafahl}
    \date{30 August,2017}

    \begin{document}
    \maketitle

    1. Use mathematical induction to prove the following statements
    $ \newline \newline $
    (a)
    $ \newline $
    $ \forall_{n} \geq 1 $, $ \sum_{i = 1}^{n} i^{3} $ = $ ( \sum_{i = 1}^{n} i )^{2} $
    $ \newline \newline $
    $ \newline $
    lemma: $ \sum_{i = 1}^{n} i $ = $ \frac{n^{2}}{2} $ + $ \frac{n}{2} $ where $ n \in N $
    $ \newline $
    Base Case: $ \sum_{i = 1}^{1} i $ = 1 = $ \frac{1^{2}}{2} $ + $ \frac{1}{2} $
    $ \newline $
    Inductive Hypothesis: $ \sum_{i = 1}^{n} i $ = $ \frac{n^{2}}{2} $ + $ \frac{n}{2} $
    $ \newline $
    Inductive Steps:
    $ \newline $
    Using induction on n
    $ \newline $
    $ \frac{(n + 1)^{2}}{2} + \frac{n + 1}{2} $ = $ \frac{n^{2}}{2} $ + $ \frac{3n}{2} $ + 1 =
    $ \frac{n^{2}}{2} + \frac{n}{2} $ + $ (n + 1) $
    $ \newline \newline $
    So therefore the lemma is true for all natural numbers n greater than 1.
    $ \newline \newline $
    With the lemma we can solve the proof easier
    $ \newline \newline $
    Base Case: For n = 1 would be $ \sum_{i = 1}^{1} i^{3} $ = $ 1^{3} $ = 1.
    $ ( \sum_{i = 1}^{1} i )^{2} $ = 1.
    $ \newline \newline $
    Inductive Hypothesis: From the base n is true for
    $ \forall_{n} \geq 1 $, $ \sum_{i = 1}^{n} i^{3} $ = $ ( \sum_{i = 1}^{n} i )^{2} $
    $ \newline \newline $
    Inductive Step: Using lemma we can re-write this to
    $ (\frac{n^{2}}{2}  +  \frac{n}{2})^{2} $ = $ \sum_{i = 1}^{n} i^{3} $
    We will first add n+1, $( \frac{(n + 1)^{2}}{2}  +  \frac{(n + 1)}{2})^{2} $ = $ \frac{n^{4}}{4} $ +
    $ \frac{3n^{3}}{2} $ + $ \frac{13n^{2}}{4} $ + 3n + 1 .
    Now we will do $ ( \frac{n^{2}}{2}  +  \frac{n}{2})^{2} $ + $ (n + 1)^{3} $ =
    $ \frac{n^{4}}{4} $ +
    $ \frac{3n^{3}}{2} $ + $ \frac{13n^{2}}{4} $ + 3n + 1 . So therfore
    $ \forall_{n} \geq 1 $, $ \sum_{i = 1}^{n} i^{3} $ = $ ( \sum_{i = 1}^{n} i )^{2} $
    is true for all natural numbers for n greater than or equal to 1.
    $ \newline \newline $
    (b) $ \forall_{n} \geq 4$, $ 2^{n} < n!$
    $ \newline \newline $
    Basis: for n = 4 $ 2^{4} $ = 16 and 4! = 24
    $ \newline \newline $
    Inductive Hypothesis: $ \forall_{n} \geq 4$, $ 2^{n} < n!$ is true for n.
    $ \newline \newline $
    Inductive Step: $ 2^{n+1} $ = $ 2^{n} * 2 $. (n + 1)! = n! * (n + 1)
    Since we know from the inductive hypothesis that $ \forall_{n} \geq 4$, $ 2^{n} < n!$
    so for this to remain true the right side and the left would need to be multiplied by the same value
    or the right would need to be multiplied by a greater value. The left is multiplied by 1 and the right is multiplied by
    (n + 1). We know that n has to be 4 or greater so we know (n + 1) > 2 so therfore $ \forall_{n} \geq 4$, $ 2^{n} < n!$
    for all natural numbers n.
    $ \newline \newline $
    2. Refer to the definition of Full Binary Tree from the notes. For a Full Binary Tree T, we
    use n(T), h(T), i(T) and `(T) to refer to number of nodes, height, number of internal nodes
    (non-leaf nodes) and number of leaves respectively. Note that the height of a tree with single
    node is 1 (not zero). Using structural induction, prove the following:
    $ \newline \newline $
    (a) For every Full Binary Tree T, n(T) $ \geq $ h(T).
    $ \newline \newline $
    Basis: For a full binary tree T with a root node and two child leaf nodes 
    $ n(T_{*}) = 3 $, and $ h(T_{*}) = 2 $ so $ n(T_{*}) \geq  h(T_{*}) $ is true.
    $ \newline \newline $
    Inductive Hypothesis: Suppose that for complete binary trees T,  n(T) $ \geq $ h(T) is true.
    $ \newline \newline $
    Recursive: Suppose we have two sub trees $ T_{1} , T_{2} $, these threes are identical full binary trees and when we combine them together with an addictional node 
    $ T_{1} + T_{2} N_{1} $ = T, $ n(T_{1,2})  \geq  h(T_{1,2}) $. First we will compare $ n(T_{1}) + n(T_{2}) + 1 \geq  h(T_{1,2}) + 1 $. $ n(T_{1}) + n(T_{2}) + 1 $ = n(T)
    and h(T) = $ h(T_{1,2}) $ + 1. We can reduce it down to $ 2n(T_{1}) + 1 \geq  h(T_{1}) + 1 $ .... $ 2n(T_{1}) \geq  h(T_{1}) $. By the IH we know that $ n(T_{1}) \geq  h(T_{1}) $
    is true so $ 2n(T_{1}) \geq  h(T_{1}) $ must also be true so therfore n(T) $ \geq $ h(T) is true for all full binary trees T.
    $ \newline \newline $
    (b) For every Full Binary Tree T, i(T) $ \ge $ h(T) - 1
    $ \newline \newline $
    Basis: For a full binary tree $ T_{*} $ with a single root node and two child leaf nodes. 
    $ i(T_{*}) $ = 1 and $ h(T_{*}) - 2 $ = 1 so therfore $ i(T_{*})  \ge  h(T_{*}) - 1 $
    $ \newline \newline $
    Inductive Hypothesis: Assume that i(T) $ \ge $ h(T) - 1 is true.
    $ \newline \newline $
    Recursive: We will take two idential full binary trees $ T_{1,2} $. If we  
    combined them with a single node to connect them into a new full binary tree T
    then to find all the internal nodes of T we would do $ i(T_{1}) + i(T_{2}) + 1 \ge h(T) - 1 $. 
    The height of T is one more than $ T_{1} $ so $ i(T_{1}) + i(T_{2}) + 1 \ge h(T_{1}) $. This can also 
    be reduced down to $ i(2T_{1}) + 1 \ge h(T_{1}) - 1 $. By the IH $ i(T_{1}) + 1 \ge h(T_{1}) - 1 $ 
    so $ i(2T_{1}) + 1 \ge h(T_{1}) - 1 $ is also true and therfore i(T) $ \ge $ h(T) - 1 is true
    for all complete binary trees T.
    $ \newline \newline $
    (c) For every Full binary Tree T, $ l(T) = (n(T) + 1)/2 $
    $ \newline \newline $
    Basis: for a full binary tree $T^{2} $ with a single root node and two children has 2 leaf nodes and 3 alltogether.
    so (3 + 1)/2 = 2, so therefore $ l(T^{3}) = (n(T^{3}) + 1)/2 $ is true for the basis.
    $ \newline \newline $
    Inductive Hypothesis: We assume $ l(T) = (n(T) + 1)/2 $ is true.
    $ \newline \newline $
    Recursive Step: We will take two idential full binary trees $ T_{1,2} $. If we  
    combined them with a single node to connect them into a new full binary tree T.
    By the IH $ l(T_{1,2}^{3}) = (n(T_{1,2}^{3}) + 1)/2 $. To find the number of leaves
    for T we can add these together so $ \frac{n(T_{1}) + 1}{2} $ + $ \frac{n(T_{2}) + 1}{2} $.
    We can simpllify this to because $ T_{1} $ = $ T_{2} $ , $ n(T_{1}) + 1 $ = l(T). 
    Now me must prove this is correct. We know that $ l(T) = (n(T) + 1)/2 $ so we will 
    expand n(T) in terms of $ T_{1} $ so $ l(T) = (2n(T_{1}) + 1 + 1)/2 $ ... l(T) = $ n(T_{1}) + 1 $.
    So because we were able to prove for T with two sub trees therefore $ l(T) = (n(T) + 1)/2 $ is true for 
    all full binary trees T.
    $ \newline \newline $
    3. Let a be an array of size n indexed by 0, 1, ... n − 1. Consider the following code (inner loop
    of Bubble Sort)
    $ \newline \newline $
    for i in the range [0, n-2]
    $ \newline $
        if (a[i] $>$ a[i+1])
        $ \newline $
            swap (a[i], a[i+1]); //swap the values of a[i] and a[i+1]

    $ \newline \newline $
    Show the following using induction :“ At the start of ith iteration of the loop, a[i] is the
    maximum among a[0], a[1], ... a[i].
    $ \newline \newline $
    Basis: for a[0] there are no values index in a below 0 so therfore a[i] where i = 0 
    the start of the iteration of i is the maximum.
    $ \newline \newline $
    inductive Hypothesis: Assume that a[i] is the greatest max value during the iteration 
    and that m(a[i]) is either true or false if a[i] is greatest during the ith iteration.
    $ \newline \newline $
    Inductive Step: By the IH assume that m(a[i]) is true so. If during the ith interation 
    a[i] $>$ a[i+1] then a[i+1] = a[i] in that case for m(a[i+1]) a[i+1] $>$ a[i] and m(a[i]) is true by
    the inductive hypothesis so therefore for that that m(a[i+1]) is true. The other
    case is that a[i+1] $>$ a[i] during the ith iteration. In that case its the same
    where m(a[i]) is already true so starting at the i+1 iteration we already know a[i]
    is greater than anything before it and the value after a[i], is greater a[i+1] $>$ a[i]
    so therefore m(a[i]) or that a[i] is the greatest max for a[0],a[1],...a[i] for all
    natural numbers i where $ 0 \leq i \leq n-1 $.
    $ \newline \newline $
    4. Let a be an array of size n indexed by 0, 1, ... n - 1. Now consider Bubble sort pseudo code.
    $ \newline \newline $
    for j in the range [0, n-1]
    $ \newline \newline $
    for i in the range [0, n-j-2]
    $ \newline \newline $
    if a[i] $>$ a[i+1]
    $ \newline \newline $
    swap(a[i], a[i+1])
    $ \newline \newline $
    Use mathematical induction to show: At the start of jth iteration of outer loop the following
    conditions hold:
    $ \newline \newline $
    a[n - j] $ \leq $ a[n - j + 1] $ \leq $ a[n - 1] 
    $ \newline \newline $
    a[n - j], a[n - j + 1], ... a[n - 1] are the j largest elements of the array
    $ \newline \newline $
    Basis: Using induction on j we assume j = 0, a[n + 1] or a[n] does not exist so it is true
    because the implication is false therefore the expression is true.
    $ \newline \newline $
    Inductive Hypothesis: With the conditions held in the basis we will assume for iteration j
    that the properties.
    $ \newline \newline $
    a[n - j] $ \leq $ a[n - j + 1] $ \leq $ a[n - 1] 
    $ \newline \newline $
    a[n - j], a[n - j + 1], ... a[n - 1] are the j largest elements of the array
    $ \newline \newline $
    Inductive Step: We need to prove the property for
    a[n - j - 1] $ \leq $ a[n - j] $ \leq $ a[n - 1] and
    a[n - j - 1], a[n - j], ... a[n - 1] are the j largest elements of the array.
    By the IH we already know that $ a[n - j] \leq a[n - 1] $. For the j+1 iteration and by 
    the def of problem 3 that when the iner loop reaches iteration n-j-1, n-j-1 will be the greatest value 
    of every index before it. The for loop as not altered this value before this point so it was the greatest
    value at the beginning of the iteraiton. We also know from the IH that a[n - j], a[n - j + 1], and a[n - 1]
    are the greatest values so we know that a[n - j - 1] $ \leq $ a[n - j]. We know that a[n - j - 1] is
    greater than anything before it and we know that $ a[n - j - 1] \leq a[n - j] \leq a[n - 1] $ so we know that
    a[n - j - 1], a[n - j], a[n - 1] are the largets elements of the array. Therefore 
    $ \newline \newline $
    a[n - j] $ \leq $ a[n - j + 1] $ \leq $ a[n - 1] 
    $ \newline \newline $
    a[n - j], a[n - j + 1], ... a[n - 1] are the j largest elements of the array
    $ \newline \newline $
    are true for all iterations of j.
    $ \newline \newline $
    5. This proof is incorrect because it assumes that $ c_{m+1} $ = $ \forall_{a} \in S_{2} $. 
    


    \end{document}
